%\section{Introduction}
\section{Background} 
Review of literature and background study \\

\textbf{[In case of internship - Company Background]}
Describe the company where you completed your internship. Include information such as the company’s history, mission, vision, key products or services, and its position in the industry. Highlight any relevant details that provide context for the project or tasks you worked on.

\section{Rational of the Study or Motivation}
This section discusses the significance and relevance of the research gap and its impact.

\section{Problem Statement}
Nowadays, Automatic license-plate recognition (ALPR) systems are built on law enforcement, smart transportation and traffic-safety surveillance infrastructures.However, due to movement of vehicles at high speed, the amount of motion blur and geometric distortion is induced into the images and there is also a significant difficulty to address non-Latin scripts, such as Bangla compared to standard english digits where the structural complexity of the numerals and the contacts between the characters (Matra) place a greater thinking burden on the observer[1]. Even though few researchers have addressed a solution, there remains some technical gaps.Even the previous studies focus on restoration of images using GFPGAN which specially targets low-resolution motion blur but not the high frequency motion blur[1].Moveover,Other studies by  Wang et al. [2], uses YOLOv5 to ensure environment interference and lilted plates are restored which do not include a fast moving car’s licence plate distortion fixing Where as 
low-light enhancement studies that use URetinex-Net [3] and U-Net or YOLOv10 [4], [5]
Handles atmospheric noises illumination changes fails when there is motion-induced blur for more demanding character level integrity. So to overcome this gap our study will propose a  powerful multi-stage deep learning architecture which will handle the motion-blurred and geometrically distorted Bengali license plates.Our suggested design uses the YOLOv8 to identify the object accurately, DeblurGANv2 to evaluate the high-speed image enhancement specifics, and a Bangla-trained OCR model to consider the script-specifics.This system offers a real world reliability which involves heavy traffic and environments aimed at preventing crime,thus providing for Bangladesh's intelligent transportation infrastructure an adaptable approach.


\section{Objective}
Our goal for this research is to design a multi-stage, multifaceted deep learning system which is designed to detect and recognize Bangla license plates in extreme motion blur and geometric distortion in images. This study deals with the significant decrease in performance of license plate detection for a fast moving vehicle in uncontrolled condition. For complex urban backgrounds our research uses state-of-the-art YOLOv8 architecture to make sure the detection is highly sensitive. One of the important stages of our study is to use DeblurGANv2 to reduce the loss of character-level accuracy, which is a denoising assistant to restore high-frequency texture detail in the image of a license plate when in motion. Additionally, to address the unique problems of the non-Latin character structure and matra relations that cannot be solved by the traditional models, our study intends to give a specialized optical character recognition (OCR) system which will be trained on complex Bengali script. Therefore, the ultimate goal of this study is to build an effective pipeline which significantly increases reliability and recognition accuracy in real world systems such as highway surveillance and criminal investigations. In order to improve the effectiveness of law enforcement operations and  intelligent transportation systems in Bangladesh, our research may offer an effective In order to improve intelligent transportation systems and the effectiveness of law enforcement operations in Bangladesh, this study may offer a scalable and effective way to address the academic gap between current fixing techniques and the unique script requirements of this nation.


\section{Methodology in Brief }

This research follows a practical, step-by-step methodology to design, develop, and evaluate a robust automatic Bangla license plate recognition (ALPR) system for fast-moving vehicles under real-world conditions.

\begin{enumerate}
    \item \textbf{Dataset Preparation}

    \begin{enumerate}
        \item Existing publicly available license plate datasets will be reviewed and utilized for initial model training and benchmarking.
        \item Additional Bangla license plate images will be collected from traffic surveillance footage, roadside cameras, and publicly available sources.
        \item The dataset will include images captured under challenging conditions such as:
        \begin{enumerate}
            \item High vehicle speed and motion blur
            \item Low illumination and nighttime scenarios
            \item Rain, fog, and partial occlusion
        \end{enumerate}
        \item All collected images will be annotated with:
        \begin{enumerate}
            \item License plate bounding boxes
            \item Corresponding Bangla text labels
        \end{enumerate}
        \item The dataset will be organized in a structured format to support detection, enhancement, and OCR training.
    \end{enumerate}
    \item \textbf{Preprocessing and License Plate Detection}
        \begin{enumerate}
            \item Input images or video frames will be resized and normalized for efficient processing.
            \item The YOLOv8 object detection model will be employed to localize license plate regions in each frame.
            \item Detected license plate regions will be cropped and forwarded to subsequent processing stages.
            \item Data augmentation techniques such as rotation, scaling, and brightness adjustment may be applied to improve model robustness.
        \end{enumerate}
    \item \textbf{Image Enhancement and Deblurring}
        \begin{enumerate}
            \item Cropped license plate images will be processed using DeblurGANv2 or ESRGAN to reduce motion blur and distortion.
            \item The enhancement model will aim to:
            \begin{enumerate}
                \item Restore edge sharpness
                \item Improve character visibility
                \item Increase resolution for better OCR performance
            \end{enumerate}
            \item Enhanced images will be standardized to a fixed resolution before recognition.
        \end{enumerate}
    \item \textbf{Bangla License Plate Recognition (OCR)}
        \begin{enumerate}
            \item The enhanced license plate images will be passed to a Bangla-trained OCR model for text recognition.
            \item A segmentation-free recognition approach will be used to handle complex Bangla characters.
            \item The OCR output will generate Bangla characters in digital text format.
            \item Post-processing techniques such as character validation and format checking may be applied to improve accuracy. 
        \end{enumerate}
    \item \textbf{System Integration and Real-Time Processing}
        \begin{enumerate}
            \item All modules detection, enhancement, and recognition will be integrated into a unified pipeline.
            \item The system will be optimized to support real-time or near real-time inference
            \item Processing latency and memory usage will be monitored to ensure practical deployment feasibility.
        \end{enumerate}
    \item \textbf{Evaluation and Performance Analysis}
        \begin{enumerate}
            \item The proposed system will be evaluated using quantitative metrics, including:
                \begin{enumerate}
                    \item Detection accuracy
                    \item Recognition accuracy
                    \item Precision, recall, and F1-score
                    \item Processing time per frame    
                \end{enumerate}
            \item Qualitative evaluation will be conducted through visual inspection of recognition results under different traffic conditions.
            \item The system’s robustness will be tested across variations in speed, lighting, weather, and camera angles.
        \end{enumerate}
\end{enumerate}

\section{Scopes and Challenges}

\subsection{Scope of the Research:}
This research focuses on the development of a robust, deep learning–based automatic license plate recognition (ALPR) system for Bangla license plates captured from fast-moving vehicles. The scope of the research includes:
    \begin{enumerate}
        \item Designing a multi-stage deep learning pipeline for Bangla license plate detection, enhancement, and recognition.
        \item Implementing a real-time license plate detection module using YOLOv8 for accurate localization in complex traffic environments.
        \item Integrating an image enhancement module using DeblurGANv2 or ESRGAN to reduce motion blur, distortion, and low-resolution effects.
        \item Developing or adapting a Bangla-trained Optical Character Recognition (OCR) model for recognizing non-Latin Bangla characters.
        \item Evaluating the system under challenging real-world conditions such as high-speed motion, low illumination, rain, and partial occlusion.
        \item Assessing system performance in terms of detection accuracy, recognition accuracy, processing time, and robustness.
    \end{enumerate}
 This research is primarily focused on still images or video frames captured from traffic surveillance systems, with an emphasis on practical applications in traffic monitoring, law enforcement, and intelligent transportation systems in Bangladesh.

\subsection{Challenges:}

Despite its potential impact, the proposed research faces several technical and practical challenges:
    \begin{enumerate}
        \item \textbf{Motion Blur and Image Distortion:} License plates captured from fast-moving cars often suffer from severe motion blur and geometric distortion, which significantly degrade detection and recognition accuracy, even for advanced deep learning models.
        \item \textbf{Variability in Bangla License Plate Design} Bangla license plates exhibit variations in font style, character spacing, background color, and plate condition, making consistent recognition more challenging.
        \item \textbf{Limited Bangla License Plate Datasets} There is a lack of large-scale, publicly available datasets specifically for Bangla license plates, particularly those containing blurred and distorted samples, which affects model training and generalization.
        \item \textbf{Real Time Performance Constraints:} Ensuring that the complete pipeline detection, enhancement, and OCR operates in real time or near real time is challenging due to the computational cost of image enhancement models such as DeblurGANv2 or ESRGAN.
        \item \textbf{Environmental and Lighting Variations:} Changes in illumination, weather conditions, camera angles, and occlusion from other vehicles can negatively impact system performance and robustness.
        \item \textbf{Deployment and Scalability Issues:} Deploying the system in real-world traffic environments requires careful optimization to balance accuracy, speed, and hardware limitations, especially for edge or low-resource devices.  
    \end{enumerate}




